Network Evolution

### 2021-04-28 Figs/random_network/Porous - Instability/shear_rate_homogenity.tex
Initial Overleaf Import

\input{definitions.tex}
\def\d{\text{d}}
\def\vep{\varepsilon}
\def\ep{\epsilon}
\def\la{\langle}
\def\ra{\rangle}
\def\th{\theta}
\usepackage{tikz}
\usepackage{circuitikz}
\usepackage{float}
\usepackage{hyperref}
\newcommand{\AZ}[1]{{\color{red}{Ahmad: #1}}}
\newcommand{\red}[1]{{\color{red}{#1}}}
\title{Porous Media Project}
\author{Ahmad}
\date{}

\begin{document}
\maketitle

\tableofcontents

\section{Introduction}
%
In order to study the porous media, we introduce a model that captures
flow behavior through a porous material. We consider the porous media
as a network of pores that are connected together through
connections. The schematic of our model is shown in
Fig. \ref{porous-schematic}. We use a random resistor networks on a
structured grid as our model to study the flow behavior. Each node
represents a pore and each edge represents the throats between
pores. We use simple Poiseuille flow through pipes to model fluid flow
through pipes. In each pipe with diameter $d$, length $L$, and pressure difference
of $\Delta P$ on both sides, results in a flow $Q$, where
%
\begin{align}
    \frac{\Delta P}{L}  = \frac{128}{\pi} \frac{\mu Q}{d^4} 
\end{align}
%
where $\mu$ is the viscosity of the fluid. The above relation is known
as Darcy's law.

%
\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{./Figs/porous-schematic}
  \caption{Schematic of the network model for porous media: (a) a
    sample cross section of a porous media. (b) The porous media can
    be considered as a network of pores connected through resistive
    connections. (c) The random network obtained from porous
    media. (d) Reordering the porous media network on a structured
    grid. (e) The ordered structured network used where the
    connections are assumed to be random resistors.} \label{porous-schematic}
\end{figure}
%

\section{Numerical Simulation}
%
We consider a network of tubes with a random distribution of diameters
as shown in Fig. {porous-schematic}e. Darcy's law is obtained thought
the momentum equation and relates the pressure and the flow through
the pipes. Furthermore, at each node the summation of the inflows
should add up to zero (conservation of mass). Using this a system of
equations $AP=B$ is obtained where $P$ is the vector of pressures at
nodes, $A$ is sparse where the nonzero elements are
% 
\begin{align}
  A_{ii} &= \sum_{j \text{ neighbors of } i} \frac{\pi }{128 \mu L} d_{ij}^4 \\
   A_{ij} &= 
  \begin{cases}
    -\frac{\pi }{128 \mu L} d_{ij}^4, \quad \text{if }j \text{ is
      neighbor of }i\\
    0, \quad \text{otherwise}
  \end{cases}
\end{align}
%
Note that if all the pipes have the same diameter, then the flow is
only in the horizontal pipes, and vertical pipes will carry no flow.

\subsection*{Numerical results}
%
We consider a 2d network of pipes, and a pressure difference between
the left and right nodes. We further assume different distribution of
pipe diameters and find the flow $q$ in the pipes. When the pipes are
all the same diameter, then the velocities in the pipes are either $0$
or some nonzero value. As we increase the randomness in the pipes, we
observe that the PDF of velocities starts to follow an exponential
tail trend. Interestingly, if the disorder in pipes diameter is large
enough, the PDF of velocities in the pipes, no matter what
distribution we pick follows the same distribution. 
%
\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\textwidth]{./Figs/distribution}
  \caption{PDF of normalized velocities of the pipes in a random
    network of pipes, $v$ is the velocity in the pipe, $Q$ is the
    total flux, $N$ is the number of pipes in the vertical direction,
    $\langle A\rangle$ is the average of the area of the pipe. Note that
    $N\langle A\rangle$ is the average area carrying the flow $Q$. } \label{pdf-result}
\end{figure}

\section{Analytical results}
%
\label{sec:analytical}
In order to describe the universality observed above, we introduce the
following mode. First, instead of using the structured rectangular
grid, we use a structured diagonal grid. This will simplify the
analytical model (see Fig. \ref{grid-result}).

%
\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{./Figs/grid-analytica}
  \caption{Schematic of grids: (a) The structured rectangular grid
    used in the numerical simulation. (b) The structured diagonal grid
    used in the analytical model. } \label{grid-result}
\end{figure}
%
We assume that the total flow to node $i$ at level $L$ is noted by
$Q(L,i)$. This flow is then redistributed to the neighboring nodes at
the level $L+1$. We assume that this redistribution is done by weights
$w_{ij}$ as
%
\begin{align}
  Q(L+1,j) = \sum_i w_{ij} Q(L,i) = w_{i,i+1} Q(L,i+1) + w_{i,i} Q(L,i) \label{total-flow}
\end{align}
%
where $\sum_{j} w_{ij} = 1$ (since the total outflow from node $i$ is
$\sum_j q_iw_{ij}$ and this value should add up to $q_i$). We assume
that $w_{ij}$ are drawn from a distribution $\eta(w)$. Since $\eta(w)$
is a distribution, then $\int \eta(w) dw = 1$. Furthermore, since
$\sum_jw_{ij} = 1$, we can conclude that
%
\begin{align}
  \sum_j w_{ij}=1 \to N {E}[w_{ij}] = 1 \to E[w_{ij}] = 1/N \to {\int w\eta(w) dw = 1/N}
\end{align}
%

Now we use the technique of mean field analysis for a general
distribution of $\eta(w)$ to find the distribution of $Q$ at the
layers. The values of $Q(D,i)$ are not independent for neighboring
sites; however, the mean field approximation ignores these
correlations. We have
%
\begin{align}
  P_L (Q) = \prod_{j=1}^N \left\{ \int_0^1 d w_j \eta(w_j) \int_0^{\infty} dQ_j P_{L-1}(Q_j)\right\} \times \delta \lp \sum_j w_j Q_{j} - Q \rp
\end{align}
%
The constraint that $Q$'s emanating downward should add up to one is
in the definition of $\eta(w)$. The only approximation in the above
equation is that we neglect the possible correlation between the
values of $Q$ among ancestors. If we take the Laplace transform of
the above equation, we obtain
%
\begin{align}
  \tilde P(s) & \equiv \int_0^{\infty} P(Q) e^{-Qs} dQ\\
  \xrightarrow{\int_0^{\infty} (\cdot) e^{-Qs} dQ}  P (Q) & = \prod_{j=1}^N \left\{ \int_0^1 d w_j \eta(w_j) \int_0^{\infty} dQ_j P(Q_j)\right\} \times \delta \lp \sum_j w_j Q_{j} - Q \rp \\
  \tilde P(s) & = \int_0^{\infty} \prod_{j=1}^N \left\{ \int_0^1 d w_j \eta(w_j) \int_0^{\infty} dQ_j P(Q_j)\right\} \times \delta \lp \sum_j w_j Q_{j} - Q \rp e^{-Qs}dQ \\
              & = \prod_{j=1}^N \left\{ \int_0^1 d w_j \eta(w_j) \int_0^{\infty} dQ_j e^{-\sum w_{j}Q_{j}} P(Q_j)\right\}   \\
              & = \prod_{j=1}^N \left\{ \int_0^1 d w_j \eta(w_j) \tilde P(s w_j) \right\} \\
  & = \lp  \int_0^1 d w \eta(w) \tilde P(s w)  \rp^{N}
\end{align}
%
Or to summarize
%
\begin{align}
  \boxed{\tilde P(s) = \lp  \int_0^1 d w \eta(w) \tilde P(s w)  \rp^{N}} \label{eq:main-laplace}
\end{align}
%
Note that in the above equation $N$ determines the number of
neighboring sites and in our structured diamond grid $N=2$. The above
equation is recursively converges to a distribution.  If we set
$\eta(w) = \delta(w-1/2)$, then the flow at each node has probability of
$1/2$ to move up or down. As a result the flow at a given depth $L$
will be homogeneous as $P_L(Q) = \delta (Q-\bar{Q})$ where $\bar{Q}$
is the average of input fluid flow. This case is the same as the same
diameter which results in a singular solution in the structured grid.

\subsection*{$P(Q)$ decays faster than $Q^{-n}$ for any $n$}
We first show that $P(Q)$ decays faster than any power of $Q$ for any
weight distributions. We expand the Laplace transform as
%
\begin{align}
  \tilde P(s) = 1 + \sum_j \tilde P_{j} s^{j}
\end{align}
%
Inserting this expansion into Eq. \eqref{eq:main-laplace}, we obtain
%
\begin{align}
  1 + \sum_j \tilde P_{j} s^{j}  & = \lp  \int_0^1 d w \eta(w) \lb 1 + \sum_j \tilde P_{j} (sw)^{j}\rb    \rp^{N} \\
 & = \lp  1 + \sum_j \tilde P_j \langle w^{j}\rangle  s^{j}   \rp^{N}
\end{align}
%
where $\langle w^j \rangle = \int_0^1 \eta(w) w^j dw$ is the $j$-th
moment of the weights $w$. In the above equation we can observe that
%
\begin{align}
  P_j \lp N \langle w^j\rangle -1 \rp = G(P_{j-1}, P_{j-1}, \cdots, P_1)
\end{align}
%
In the above equation $P_j$ can only diverge only if $N\langle w^j
\rangle -1$ is zero. If $w$ can only be zero and one, then $\langle
w^j\rangle =\langle w \rangle = 1/N$ and the coefficient can become
zero! Other than this special distribution where $0 \leq w \leq 1$,
then
%
\begin{align}
  & w\in [0,1] \rightarrow w>w^2>w^3> \cdots \\
  & \forall k>1 \rightarrow \langle w^k \rangle < \langle w \rangle = \frac{1}{N} 
\end{align}
%
which means that $N\langle w^j \rangle -1$ is not zero and as a result
$P_j$ is finite. $P_j$ on the other hand is the $j$-th moment or
$\langle Q^j\rangle$ and it is finite. If $P\approx Q^{-n}$, then the
$n$-th moment will not be finite i.e. $\langle Q^{n+1}\rangle =
P_{n+1}$ is not finite. We can then conclude that $P(Q)$ should go to
zero faster than any $Q^{-n}$ as $Q\to \infty$! In other words $d \log
P(Q) / d \log Q \to -\infty$ as $Q\to \infty$.

\subsection*{Exponential Decay}
%
The main governing equation for a diamond grid is 
%
\begin{align}
  \tilde P(s) = \lp  \int_0^1 d w \eta(w) \tilde P(s w)  \rp^{2} 
\end{align}
%
In order to have an approximation for $\eta(w)$ we do the following:
At each node there are two nodes where the flux is distributed by
$w_1$ and $w_2$. Lets assume that $w_1$ is uniformly distributed
between $0$ and $1$. As a result $w_2 = 1-w_1$ (since conservation of
mass $w_1 + w_2 = 1$). Now $\eta(w)$ can be found as
%
\begin{align}
  & \eta(w) = M \int_0^1 dw_1 \delta(1-w_1-w) =  M, \\
  \text{since }& \int_0^1 \eta(w) dw = 1 \to M =1 \\
  & \eta (w) = 1
\end{align}
%
Just for completeness, if there are three connections ($N=3$), then we
have
%
\begin{align}
  \eta(w) & = M \int_0^1 dw_1 \int_0^1 dw_2 \delta(1-w_1-w_2 - w)\\
          & = M(1-w)  \to \int_0^1\eta(w) dw = 1 \to M = \frac{1}{2} \\
  & \eta(w) = \frac{1}{2} (1-w)
\end{align}
%
As a result the main equation simplifies to
%
\begin{align}
  \tilde P(s) = \lp  \int_0^1 d w  \tilde P(s w)  \rp^{2} 
\end{align}
%
In order to solve the above equation, we assume $\tilde V (s) =
\sqrt{\tilde{P}(s)}$.
%
\begin{align}
  \tilde{V}(s) & = \int_0^1 dw \tilde V (sw) = \int_0^s \frac{du}{s}  \tilde{V}^{2}(u) \\
  s\tilde{V}(s) & = \int_0^s du  \tilde{V}^2(u) \\
\end{align}
%
Taking the differentiation with respect to $s$ yields
%
\begin{align}
  \tilde{V}(s) + s \frac{d \tilde{V}(s)}{d s} = \tilde{V}^2(s) \\
  \frac{d \tilde{V}}{\tilde{V}^2-\tilde{V}} = \frac{ds}{s}  \\
  \log\lp \frac{1-\tilde{V}}{\tilde{V}} \rp  = \log(s)\\
  \tilde{V}(s) = \frac{1}{1+Cs} 
\end{align}
%
We set the mean to $1$ i.e.,
%
\begin{align}
  \int_0^{\infty} Q P(Q) dQ = 1 \to \frac{d\tilde{P}}{ds}|_{s=0}  = -1\\
  \frac{d \tilde{V}}{ds}  = \frac{1}{2 \sqrt{\tilde{P}(s)}} \frac{d \tilde{P}(s)}{d s}   \to \frac{d\tilde{V}(s)}{ds}|_{s=0} =\frac{-1}{2} \\
  \frac{d\tilde{V}(s)}{ds} = \frac{-C}{(1+Cs)^{2}}|_{s=0} = -C = \frac{-1}{2}  \to C=\frac{1}{2} 
\end{align}
%
As a result, we have
%
\begin{align}
  \tilde{P}(s) = \lp\frac{1}{1+s/2} \rp^2 \\
  P(Q) = 4Q e^{-2Q}
\end{align}
%
We can then show that for a uniform distribution of $w$'s, the PDF of
$Q$'s become an exponential tail. For other distribution of $w$'s, we
similarly find a similar trend. Actually, we should that the PDF of
$Q$ should decay faster than any power law for any distribution of
$w$. Note that the only parameter that determines the the PDF of the
flow is the average flux. This shows that why we observe the universal
behaviour for pdf.

\section{Continuous Limit}
%
In the previous section (\S \ref{sec:analytical}), we used a toy model
to describe the flow behavior. The main equation connecting the flow
at different layers was Eq. \eqref{total-flow}. In this section, we
are interested to see what a continuous limit of the above model will
be. Consider Eq. \eqref{total-flow} as
%
\begin{align}
  Q(i,j+1) = w_{i-1,j} Q(i-1,j) + w_{i+1,j} Q(i+1,j) 
\end{align}
%
where we had a constraint on the weights as 
%
\begin{align}
  w_{i-1,j} + w_{i+1,j} = 1
\end{align}
%
Now, lets assume that these weights are perturbations around the a
uniform distribution such that
%
\begin{align}
  w_{i\pm 1,j}= \frac{1}{2}( 1 \pm v_{i\pm 1,j})
\end{align}
%
with th above assumption, we find that
%
\begin{align}
  Q(i,j+1) & = \frac{1}{2} \left( 1 - v_{i- 1,j}\right) Q(i-1,j) + \frac{1}{2} \left( 1 + v_{i+ 1,j}\right) Q(i+1,j)  \\
  Q + L_y \frac{\p Q}{\p y } + L^2_y \frac{\p^{2}Q}{\p y^2}  & = \frac{1}{2} (1-v + L_x \frac{\p v}{\p x}  ) \left( Q - L_x \frac{\p Q}{\p x} +   \frac{1}{2} L_x^2 \frac{\p Q^2}{\p x^2}\right) \nonumber\\
 & \qquad   +\frac{1}{2}  (1+v + L_x \frac{\p v}{\p x} )\left( Q + L_x \frac{\p Q}{\p x} + \frac{1}{2}L_x^2 \frac{\p^2 Q}{\p x^2} \right)  \\
  L_y \frac{\p Q}{\p y } & = L_x v \frac{\p Q}{\p x} + L_x \frac{\p v}{\p x} Q + \frac{L^2_x}{2} \frac{\p^{2}Q}{\p x^2}  \\
  \frac{\p Q}{\p y}    & = \frac{L_{x}}{L_{y}} \frac{\p }{\p x} \left( v Q \right)  + \frac{L^2_x}{2L_y} \frac{\p^{2}Q}{\p x^2}  
\end{align}
%
Assuming that $L_x = L_y $ (similar length scales in both directions),
we find that
%
\begin{align}
\boxed{\frac{\p Q}{\p y}  = \frac{\p}{\p x} \left( v Q\right) + D \frac{\p^{2} Q}{\p x^2}  }
\end{align}
%
which is a convection-diffusion equation.  It seems that $y$ direction
(the direction where flow propagates into the porous medium) acts
similar to time. As the flow moves inward, the $Q$ diffuses in the $x$
direction, and convects with velocity $v$.

\section{Network Evolution}
%
Any changes to the micro-structure in a porous medium, such as solute
retention during polymer flow (effectively" clogging" some pores in
the network), affects the pore-level flow and alters the bulk
behavior. In this section, we want to study the effect of such
evolution in the network. The radius of the pipes in the network, can
reduce (due to clogging) or increase (due to erosion). The rate of
such reduction (or increase) can depend on flux, or shear rate, or
velocity, or etc. We can simplify these models with all the models
with the following equation
%
\begin{align}
  \frac{dR}{dt}  \propto \pm \frac{Q}{R^{n}} 
\end{align}
%
where $n=0,2,3$ corresponds to proportionality with flow rate,
velocity, shear gradient respectively. The plus sign corresponds to
erosion, and the negative sign corresponds to the clogging. In order
to study the effect of the above equation in whole network, we first
study the effect in a two parallel/series case.

\subsection{Erosion}
%
In erosion the rate of change of radius reads as
%
\begin{align}
  \frac{d r}{dt} & = \alpha \frac{Q}{r^{n}} 
\end{align}
%
Here, note that each pipe with radius $R$ acts as a resistor, we have
Poissoulle flow equation that relates the flow $Q$ to pressure
difference $\Delta P$ as
%
\begin{align}
  Q & = \frac{-\pi r^{4}}{8\mu L}  \Delta P\\
  Q & = -C(r) \Delta P, \quad C (r) = \frac{\pi r^{4}}{8\mu L} = {\beta}{r^{4}} 
\end{align}

\subsubsection*{Two pipes in series}

We consider two pipes in series as shown in
Fig. \ref{figure:resistor-series-1}. We are interested to see what
happens two the resistors for different initial conditions? How does
the values of the resistors evolve? What happens to the pressure
distribution at the note in between?
%
\begin{figure}[h]
  \begin{center}
    \begin{circuitikz}\draw
      (0,0) node[anchor=east] {$P_{l}$} to [short,o-] (0.1,0)
       to[R=$1/C_{1}$,-*] (2,0)  node [anchor=north] {$P_m$}
       to[R=$1/C_2$] (4,0)
       to[short,-o] (4.1,0) node [anchor=west] {$P_r$}; % The resistor
    \end{circuitikz}
    \caption{Two resistors at series}\label{figure:resistor-series-1}
  \end{center}
\end{figure}
%

We first simulate the results numerically to see what happens. The
results are shown in Fig. \ref{toy-series}. As it can be seen, for any
initial conditions the pressure at the middle point, converges to
$(P_l + P_{r})/2$ which basically means that the flow is homogenized.
%
\begin{figure}[!h]
  \centerline{\includegraphics[width=1\textwidth]{./Figs/toy-model-series}}
\caption{Toy model for two pipes in series. In each iteration, we
  change the radius of the pipe as $r_{\text{new}} = r_{\text{old}} +
  \alpha Q/r_{\text{old}}^n$. For all different cases of $n$ the
  middle pressure converges to $(P_l + P_r)/2$.}
\label{toy-series}
\end{figure}  
%

We now show why this convergence happens. If the two pipes are in
series, then $Q$ is always the same in these two pipes. Therefore
%
%
\begin{align}
  \frac{d r}{dt} & = \alpha \frac{Q}{r^{n}} \\
  r^{n}{d r} & = \alpha {Q} dt \\
  \frac{1}{n+1} r^{n+1} -  \frac{1}{n+1} r_0^{n+1} & = \alpha Q t\\
  r & = \lp (n+1)\alpha Q t + r_0^{n+1}\rp ^{\frac{1}{n+1}} \\
  C_i & = \beta \lp (n+1)\alpha Q t + r_{i,0}^{n+1}\rp ^{\frac{4}{n+1}}
\end{align}
%
So the pressure at the middle point becomes as 
%
\begin{align}
  Q &= \frac{1}{1/C_{1} + 1/C_2}   \Delta P = \text{const.}, \\
  P_m - P_l &= \frac{Q}{C_{1}}   \\
  \frac{P_{m}-P_l}{P_r - P_l} & = \frac{1/C_{1}}{1/C_{1} + 1/C_2} = \frac{C_{2}}{C_{1} + C_2}  
\end{align}
%
Combining the results we find that
%
\begin{align}
  &\boxed{\frac{P_{m}-P_l}{P_r - P_l} = \frac{\beta \lp (n+1)\alpha Q t + r_{2,0}^{n+1}\rp ^{\frac{4}{n+1}}}{\beta \lp (n+1)\alpha Q t + r_{1,0}^{n+1}\rp ^{\frac{4}{n+1}} + \beta \lp (n+1)\alpha Q t + r_{2,0}^{n+1}\rp ^{\frac{4}{n+1}}} } \\
  &\lim_{t\to\infty} \frac{P_{m}-P_l}{P_r - P_l}  = \frac{1}{2} 
\end{align}
%
Note that for any power of $n$ the pressure in the middle point as $t\to\infty$
converges to $(P_l+P_r)/2$.

\newpage
\subsubsection*{Two Parallel Pipes Case}
%
Now assume two pipes in parallel as shown in
Fig. \ref{figure:resistor-series-par}. We first simulate the results
to see what happens for different initial conditions. The results for
different initial values of $r_1,r_2$ are shown in Fig. \ref{toy-par}.
%
\begin{figure}[!h]
  \begin{center}
    \begin{circuitikz}
      \draw
      (-0.5,0) node[anchor=east] {$P_{l}$} to [short] (0.1,0)
      -- (0.1,0.5) 
       to[R=$1/C_1$] (2,0.5)  -- (2,0) to [short] (2.5,0) node
       [anchor=west] {$P_r$};
      \draw
      (0.1,0)  --(0.1,-0.5) 
       to[R=$1/C_2$] (2,-0.5)  -- (2,0);
    \end{circuitikz} 
    \caption{Two resistors parallel together} \label{figure:resistor-series-par}
  \end{center}
\end{figure}
%

\begin{figure}[!h]
  \centerline{\includegraphics[width=1\textwidth]{./Figs/toy-model-par}}
  \caption{Toy model for two pipes in parallel. In each iteration, we
    change the radius of the pipe as
    $r_{\text{new}} = r_{\text{old}} + \alpha
    Q_{i}/r_{\text{old}}^n$. When $n<3$, all of the flow eventually
    passes from either top or bottom pipe. When $n=3$, the flow
    proportion remains the same. When $n>3$, the flow homogenizes
    where half of the total flow i.e. $Q/2$ passes from top and the
    other half from bottom pipe.}
\label{toy-par}
\end{figure}

Now, we start to model to see why this happens. Initially, we note
that
%
\begin{align}
  Q_1 = C_1 \Delta P, \quad \text{ and }\quad Q_2 = C_2 \Delta P \\
  Q = (C_1 + C_2) \Delta P
\end{align}
%
We change the radius of each pipe according to its flow, as a result,
we have
%
\begin{align}
  \frac{dr_{1}}{dt}  = \alpha \frac{Q_{1}}{r_{1}^n}, \qquad  \frac{dr_{2}}{dt}  = \alpha \frac{Q_{2}}{r_{2}^n} \\
  r_1^n \frac{dr_{1}}{dt}  + r_2^n \frac{dr_{2}}{dt}  = \alpha (Q_1 + Q_2) = \alpha Q  
\end{align}
%
On the other hand, we know that
%
\begin{align}
  \Delta P = \frac{Q_{1}}{C_{1}}  = \frac{Q_{2}}{C_{2}} \\
  \Delta P = \frac{Q_{1}}{\beta r_1^{4}}  = \frac{Q_{2}}{\beta r_2^{4}}   \\
  \frac{dr_{i}}{dt}  = \alpha \frac{Q_{i}}{r^{n}_{i}} = \frac{\alpha \beta \Delta P}{r^{n-4}_{i}}   
\end{align}
%

>>>>

\input{definitions.tex}
\def\d{\text{d}}
\def\vep{\varepsilon}
\def\ep{\epsilon}
\def\la{\langle}
\def\ra{\rangle}
\def\th{\theta}
\usepackage{tikz}
\usepackage{circuitikz}

\title{Network Evolution}
\author{Ahmad}
%-------------------------------------------------------------------
\begin{document}          
\maketitle

In this report, we study the erosion or clogging in a porous media. We consider the porous material as a network of pores connected through

\section{Toy Model 1 - Erosion }
%
In each pipe with radius $R$ which acts as a resistor, we have
Poissoulle flow equation that relates the flow $Q$ to pressure
difference $\Delta P$ as
%
\begin{align}
  Q & = \frac{-\pi r^{4}}{8\mu L}  \Delta P\\
  Q & = -C(r) \Delta P, \quad C (r) = \frac{\pi r^{4}}{8\mu L} = {\beta}{r^{4}} 
\end{align}
%
We now change the radius of the pipes according to
%
\begin{align}
  \frac{d r}{dt} & = \alpha \frac{Q}{r^{n}} 
\end{align}
%

\subsection{Two in Series}
%
Now assume two pipes back to back as shown in Fig. \ref{figure:resistor-series-1}.
%
\begin{figure}[h]
  \begin{center}
    \begin{circuitikz}\draw
      (0,0) node[anchor=east] {$P_{l}$} to [short,o-] (0.1,0)
       to[R=$1/C_{1}$,-*] (2,0)  node [anchor=north] {$P_m$}
       to[R=$1/C_2$] (4,0)
       to[short,-o] (4.1,0) node [anchor=west] {$P_r$}; % The resistor
    \end{circuitikz}
    \caption{Two resistors at series}\label{figure:resistor-series-1}
  \end{center}
\end{figure}
%
If the two pipes are in series, then $Q$ is always a
constant. Therefore
%
%
\begin{align}
  \frac{d r}{dt} & = \alpha \frac{Q}{r^{n}} \\
  r^{n}{d r} & = \alpha {Q} dt \\
  \frac{1}{n+1} r^{n+1} -  \frac{1}{n+1} r_0^{n+1} & = \alpha Q t\\
  r & = \lp (n+1)\alpha Q t + r_0^{n+1}\rp ^{\frac{1}{n+1}} \\
  C_i & = \beta \lp (n+1)\alpha Q t + r_{i,0}^{n+1}\rp ^{\frac{4}{n+1}}
\end{align}
%
So the pressure at the middle point becomes as 
%
\begin{align}
  Q &= \frac{1}{1/C_{1} + 1/C_2}   \Delta P = \text{const.}, \\
  P_m - P_l &= \frac{Q}{C_{1}}   \\
  P_r - P_l &= \frac{1}{1/C_{1} + 1/C_2} Q   \\
  \frac{P_{m}-P_l}{P_r - P_l} & = \frac{1/C_{1}}{1/C_{1} + 1/C_2} = \frac{C_{2}}{C_{1} + C_2}  
\end{align}
%
Combining the results we find that
%
\begin{align}
  &\boxed{\frac{P_{m}-P_l}{P_r - P_l} = \frac{\beta \lp (n+1)\alpha Q t + r_{2,0}^{n+1}\rp ^{\frac{4}{n+1}}}{\beta \lp (n+1)\alpha Q t + r_{1,0}^{n+1}\rp ^{\frac{4}{n+1}} + \beta \lp (n+1)\alpha Q t + r_{2,0}^{n+1}\rp ^{\frac{4}{n+1}}} } \\
  &\lim_{t\to\infty} \frac{P_{m}-P_l}{P_r - P_l}  = \frac{1}{2} 
\end{align}
%
Note that if $n+1<4$ (i.e. $n<3$) then the growth is much faster, and
the limit is reached much faster; however, if $n\geq 3$ the growth
becomes slower. The result of simulation is shown in the Fig.

\begin{figure}[h]
  \centerline{\includegraphics[width=1\textwidth]{./Figs/toy-model-series}}
\caption{Toy model for two pipes in seris. In each iteration, we
  change the radius of the pipe as $r_{\text{new}} = r_{\text{old}} +  \alpha Q/r_{\text{old}}^n$}
\label{toy-series}
\end{figure}

\newpage
\subsection{Two Parallel Pipes Case}
%
Now assume two pipes in parallel as shown in Fig. \ref{figure:resistor-series-par}.
%
\begin{figure}[ht]
  \begin{center}
    \begin{circuitikz}
      \draw
      (-0.5,0) node[anchor=east] {$P_{l}$} to [short] (0.1,0)
      -- (0.1,0.5) 
       to[R=$1/C_1$] (2,0.5)  -- (2,0) to [short] (2.5,0) node
       [anchor=west] {$P_r$};
      \draw
      (0.1,0)  --(0.1,-0.5) 
       to[R=$1/C_2$] (2,-0.5)  -- (2,0);
    \end{circuitikz} 
    \caption{Two resistors parallel together} \label{figure:resistor-series-par}
  \end{center}
\end{figure}
%
In this case, first we have
%
\begin{align}
  Q_1 = C_1 \Delta P, \quad \text{ and }\quad Q_2 = C_2 \Delta P \\
  Q = (C_1 + C_2) \Delta P
\end{align}
%
We change the radius of each pipe according to its flow, as a result,
we have
%
\begin{align}
  \frac{dr_{1}}{dt}  = \alpha \frac{Q_{1}}{r_{1}^n}, \qquad  \frac{dr_{2}}{dt}  = \alpha \frac{Q_{2}}{r_{2}^n} \\
  r_1^n \frac{dr_{1}}{dt}  + r_2^n \frac{dr_{2}}{dt}  = \alpha (Q_1 + Q_2) = \alpha Q  
\end{align}
%
On the other hand, we know that
%
\begin{align}
  \Delta P = \frac{Q_{1}}{C_{1}}  = \frac{Q_{2}}{C_{2}} \\
  \Delta P = \frac{Q_{1}}{\beta r_1^{4}}  = \frac{Q_{2}}{\beta r_2^{4}}   \\
  \frac{dr_{i}}{dt}  = \alpha \frac{Q_{i}}{r^{n}_{i}} = \frac{\alpha \beta \Delta P}{r^{n-4}_{i}}   
\end{align}
%

We have
%
\begin{align}
  &r_{1}^{n-4} \frac{dr_{1}}{dt} = r_2^{n-4}\frac{dr_{2}}{dt}\\
  &\text{if }n=3, \quad \to \log(\frac{r_{1}}{r_{1,0}}) = \log(\frac{r_{2}}{r_{2,0}}), \to r_1 = \gamma r_2 \\
  & \text{if }n\neq 3, \quad r_1^{n-3} - r_2^{n-3} = \gamma
\end{align}
%
where $\gamma$ is a constant. In the above equation, without loss of
generality, we can assume that $r_{1}>r_2$ and $\gamma$ is
positive. Now, looking at the other part of the equation, we have
%
\begin{align}
  \frac{dr_{2}}{dt}  & = \alpha \frac{Q_2}{r^4_{2}} =  \frac{\alpha \beta \Delta P}{r^{n-4}_{2}} \\
                     & = \frac{\alpha \beta }{r^{n-4}_{2}} \frac{Q}{\beta (r_{2}^{4} + r_1^4)} = \frac{\alpha Q}{r^{n}_{2} + r_{2}^{n-4} r_1^{4}} \\
                     & = \frac{\alpha Q}{r^{n}_{2} + r_{2}^{n-4} \lp r_2^{n-3} + \gamma \rp^{\frac{4}{n-3} }}\\
  \lb r^{n}_{2} + r_{2}^{n-4} \lp r_2^{n-3} + \gamma \rp^{\frac{4}{n-3} }\rb dr_2 & = \alpha Q dt\\
  \frac{1}{n+1} \lb r_2^{n+1} + \lp r_2^{n-3} + \gamma\rp^{\frac{n+1}{n-3} }\rb  & = \alpha Qt + C\\
  r_2^{n+1} + \lp r_2^{n-3} + \gamma\rp^{\frac{n+1}{n-3} }  & = (n+1)\lp \alpha Qt + C\rp \\
  r_2^{n+1} + r_{1}^{n+1} = At+B
\end{align}
%
So the two set of equations summarizes to
%
\begin{align}
  r_1^{n-3} - r_2^{n-3} &= \gamma \\
  r_1^{n+1} + r_2^{n+1} &= At + B
\end{align}
%
The plot for the solution to the above equations for different values
of $n\neq 3$ and also the separate solution for $n=3$ is shown in the
following figure.  
%

\begin{figure}[h]
  \centerline{\includegraphics[width=0.8\textwidth]{./Figs/par_resistors}}
  \caption{Analytical solution for $r_1$ and $r_2$ for parallel
    resistors. For different $n$ the solution changes. Specifically,
    if $n<3$, then one of the diameters stops varying while the other
    one increases in size. If $n=3$, the both grow together, and if
    $n>3$ then they both increase in size. }
\label{anal-par}
\end{figure}

The result for numerical simulation is shown in Fig. \ref{toy-par}.

\begin{figure}[h]
  \centerline{\includegraphics[width=1\textwidth]{./Figs/toy-model-par}}
\caption{Toy model for two pipes in parallel. In each iteration, we
  change the radius of the pipe as $r_{\text{new}} = r_{\text{old}} +  \alpha Q_{i}/r_{\text{old}}^n$}
\label{toy-par}
\end{figure}

Another way of seeing this and also generalize the results is to study the change of relative radius $\tilde{r}_1 = \frac{r_1}{r_1+r_2}$ under the assumptions 
\begin{equation}
    C_i = r_i^m, ~ \frac{dr_i}{dt}= \alpha \frac{Q_i}{r_i^n}.
\end{equation}
It can be easily derived that 
\begin{equation}
    \frac{d\tilde{r_1}}{dt} \propto \tilde{r}_1 (1-\tilde{r}_1) \frac{r_1^{m-n-1}-r_2^{m-n-1}}{r_1^m+r_2^m},
\end{equation}
which has  a fixed point at either 1.  $\tilde{r_1} = 0$ or $1$ or 2.$r_1=r_2$. The first case is the channeling and the second case is the uniform. It can also be shown that if $m-n-1>0$ the system ends at channelling, otherwise uniform except when $m-n-1=0$ it will end up with no change of the relative radius.
